---
title: "Four_figures_panels"
author: "Josefina Lascano Maillard"
date: "2026-09-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Diet effect on WT vs KO mice

In this analysis, we want to compare the weigths in KO vs WT mice
```{r load mice data}
getwd()
mice_data <- read.table("course_datasets/mice_data_mod.csv", header=TRUE, sep=",") # define classes for columns
mice_data$genotype <- factor(mice_data$genotype, levels=c("WT", "KO"))
mice_data$diet <- factor(mice_data$diet, levels=c("CHOW", "HFD"))



```

## Realtionship between respiratory rate, weight, genotype and diet

We visualize the relationship between variables of interest using scatter and box plots.

```{r mice variables, echo=FALSE, fig.width=8, fig.height=6, fig.align="center"}
# 1) Make a multi-panel figure with the four graphics on one page, exporting the figure to a 
# we want to make 4 plots on the same panel -> 2 rows and 2 columns

par(mfrow=c(2,2)) # arrange subsequent plots in 2 rows x 2 columns

# Plot 1
plot(mice_data$weight,mice_data$respiratoryRate, 
     pch=19,
     main="Respiratory Rate vs Weight in Mice",
     xlab="Weight [g]", ylab="Respiratory Rate [bpm]",
     col=c("orange", "blue")[mice_data$genotype]
)

legend("bottomright",
       legend=levels(mice_data$genotype),
       col=c("orange","blue"),
       pch=19)

abline(lm(mice_data$respiratoryRate ~ mice_data$weight),
       col="black", lwd=1.5)

# Plot 2
boxplot(weight ~ genotype, data= mice_data,
        col=c("orange", "blue"),
        main="Mouse Weight by Genotype"
)
points(weight ~ genotype, data= mice_data)


# Plot 3
plot(mice_data$weight,mice_data$respiratoryRate, 
     pch=19,
     main="Respiratory Rate vs Weight in Mice",
     xlab="Weight [g]", ylab="Respiratory Rate [bpm]",
     col=c("darkred", "goldenrod")[mice_data$diet]
)

legend("bottomright",
       legend=levels(mice_data$diet),
       col=c("darkred","goldenrod"),
       pch=19)

abline(lm(mice_data$respiratoryRate ~ mice_data$weight),
       col="black", lwd=1.5)

# Plot4
boxplot(weight ~ diet, data= mice_data,
        col=c("darkred", "goldenrod"),
        main="Mouse Weight by Diet"
)
points(weight ~ diet, data= mice_data)


```


```{r create pdf, echo=FALSE, message=FALSE}
# 1) Make a multi-panel figure with the four graphics on one page, exporting the figure to a pdf file.
#    Set width and height arguments in the call to png() to make it look nice.

pdf("mice_data_plots_by_genotype.pdf", width=7, height=7, paper="a4") 

# we want to make 4 plots on the same panel -> 2 rows and 2 columns
par(mfrow=c(2,2))

# Plot 1
plot(mice_data$weight,mice_data$respiratoryRate, 
     pch=19,
     main="Respiratory Rate vs Weight in Mice",
     xlab="Weight [g]", ylab="Respiratory Rate [bpm]",
     col=c("orange", "blue")[mice_data$genotype]
)

legend("bottomright",
       legend=levels(mice_data$genotype),
       col=c("orange","blue"),
       pch=19)

abline(lm(mice_data$respiratoryRate ~ mice_data$weight),
       col="black", lwd=1.5)

# Plot 2
boxplot(weight ~ genotype, data= mice_data,
        col=c("orange", "blue"),
        main="Mouse Weight by Genotype"
)
points(weight ~ genotype, data= mice_data)


# Plot 3
plot(mice_data$weight,mice_data$respiratoryRate, 
     pch=19,
     main="Respiratory Rate vs Weight in Mice",
     xlab="Weight [g]", ylab="Respiratory Rate [bpm]",
     col=c("darkred", "goldenrod")[mice_data$diet]
)

legend("bottomright",
       legend=levels(mice_data$diet),
       col=c("darkred","goldenrod"),
       pch=19)

abline(lm(mice_data$respiratoryRate ~ mice_data$weight),
       col="black", lwd=1.5)

# Plot4
boxplot(weight ~ diet, data= mice_data,
        col=c("darkred", "goldenrod"),
        main="Mouse Weight by Diet"
)
points(weight ~ diet, data= mice_data)
dev.off()

#invisible(dev.off())
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
